NumPy---Array Programming in Python

### 2020-02-02 body.tex
Minor edits

% KJM: I moved this here to get the placement right 
% RG: minor comment on the example code: would be better to not copy IPython
%     oddities but align output on brackets and In/Out on ":"
\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{static/sketches/array-concepts}   
  \caption{\textbf{Fundamental Array Concepts.}
    \textbf{a,} The NumPy array data structure and its associated metadata fields.
    \textbf{b,} Indexing an array with various types of arguments.
    \textbf{c,} Indexing with arrays, which broadcast the indexing arguments before performing the lookup.
    \textbf{d,} Vectorization efficiently applies operations to groups of elements
    \textbf{e,} Broadcasting in scalar addition, and in the division of two-dimensional arrays.
    \textbf{f,} Reduction operations act along one or more axes. In this
    example, a three-dimensional array is shown to be summed along various single
    axes to produce two-dimensional results, or along two axes consecutively to
    produce a one-dimensional result.
    \textbf{g,} Example NumPy code, illustrating some of these concepts.
   }
  \label{fig:array-concepts}
\end{figure*}

@@

% RG: (to discuss) this figure plus associated text says that the layers about
%     Foundation build on top of that whole foundation. In reality that's not
%     true - everything builds on NumPy, almost everything on SciPy, and very
%     little on Matplotlib/IPython/Jupyter. Perhaps "interactive environment"
%     should get its own section instead.
%     Also, I think Pandas can be added next to statsmodels as statistics. That
%     avoids explaining dataframes; now it looks like a serious omission.
%     I'd also be in favor of adding Spyder in the text; scientific computing
%     beyond exploratory work is better done in an IDE.
\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{static/ecosystem}
  \caption{\textbf{Scientific Python Ecosystem.}
   This figure highlights just a few of the essential tools and libraries in the vast
   scientific Python ecosystem.  Most of the ecosystem builds on NumPy arrays to
   provide a rich and powerful collection of array-aware functionality.
  }
  \label{fig:ecosystem}
\end{figure}

@@

NumPy has subsequently seen prolific adoption, with millions of users,
first from research and later also industry.
It spurred the development of the larger scientific Python ecosystem and
heralded the current era of wide-spread use of Python for scientific computing.
The reasons for its popularity are varied and complex, but include the fact that
NumPy (a) is built on top of Python,
(b) has a strong culture of collaboration and good software development practices, and
(c) provides a cutting-edge array programming environment for scientific data analysis.

@@

% RG: this first sentence is duplicate with (b) three paragraphs up - on
%     purpose or not?
NumPy is a community of practice with a strong \emph{culture} of
employing software engineering practice to improve collaboration and
reduce error \cite{millman2014developing}.  This culture is not only
adopted by leaders in the project, but also enthusiastically taught to
newcomers. The NumPy team was early in adopting distributed revision
control and code review to improve collaboration on code, and
continuous testing that runs a large battery of automated tests for
every proposed change to NumPy.  The project has comprehensive,
high-quality documentation, integrated with the source
code \cite{vanderwalt2008scipy,harrington2008scipy,harrington2009scipy}.

>>>>

% KJM: I moved this here to get the placement right 
% RG: minor comment on the example code: would be better to not copy IPython
%     oddities but align output on brackets and In/Out on ":"
% KJM: agreed
% FIXME we should figure out if we want to keep working on the figures in this format or not
\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{static/sketches/array-concepts}   
  \caption{\textbf{Fundamental Array Concepts.}
    \textbf{a,} The NumPy array data structure and its associated metadata fields.
    \textbf{b,} Indexing an array with various types of arguments.
    \textbf{c,} Indexing with arrays, which broadcast the indexing arguments before performing the lookup.
    \textbf{d,} Vectorization efficiently applies operations to groups of elements
    \textbf{e,} Broadcasting in scalar addition, and in the division of two-dimensional arrays.
    \textbf{f,} Reduction operations act along one or more axes. In this
    example, a three-dimensional array is shown to be summed along various single
    axes to produce two-dimensional results, or along two axes consecutively to
    produce a one-dimensional result.
    \textbf{g,} Example NumPy code, illustrating some of these concepts.
   }
  \label{fig:array-concepts}
\end{figure*}

@@

% RG: (to discuss) this figure plus associated text says that the layers about
%     Foundation build on top of that whole foundation. In reality that's not
%     true - everything builds on NumPy, almost everything on SciPy, and very
%     little on Matplotlib/IPython/Jupyter. Perhaps "interactive environment"
%     should get its own section instead.
%     Also, I think Pandas can be added next to statsmodels as statistics. That
%     avoids explaining dataframes; now it looks like a serious omission.
%     I'd also be in favor of adding Spyder in the text; scientific computing
%     beyond exploratory work is better done in an IDE.
% KJM: agreed
% FIXME we should figure out if we want to keep working on the figures in this format or not
\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{static/ecosystem}
  \caption{\textbf{Scientific Python Ecosystem.}
   This figure highlights just a few of the essential tools and libraries in the vast
   scientific Python ecosystem.  Most of the ecosystem builds on NumPy arrays to
   provide a rich and powerful collection of array-aware functionality.
  }
  \label{fig:ecosystem}
\end{figure}

@@

NumPy has subsequently seen prolific adoption, with millions of users,
first from research and later also industry.
It spurred the development of the larger scientific Python ecosystem and
heralded the current era of wide-spread use of Python for scientific computing.
The reasons for its popularity are varied and complex, but include the fact that
NumPy (a) is built on top of Python,
(b) has a culture of collaboration and good software development practices, and
(c) provides a cutting-edge array programming environment for scientific data analysis.

@@

% RG: this first sentence is duplicate with (b) three paragraphs up - on
%     purpose or not?
% KJM: yes, sort of, (a) corresponds to the two para above
% (b) corresponds to the next two para and
% (c) corresponds to the last two (or three) para
NumPy is a community of practice with a strong \emph{culture} of
employing software engineering practice to improve collaboration and
reduce error \cite{millman2014developing}.  This culture is not only
adopted by leaders in the project, but also enthusiastically taught to
newcomers. The NumPy team was early in adopting distributed revision
control and code review to improve collaboration on code, and
continuous testing that runs a large battery of automated tests for
every proposed change to NumPy.  The project has comprehensive,
high-quality documentation, integrated with the source
code \cite{vanderwalt2008scipy,harrington2008scipy,harrington2009scipy}.
